---
title: "Duckweed Species Composition Presence/Absence"
output: html_notebook
author: Lillian Leak
project: final project for Biostats
---

and so it begins.
This data is presence/absence predicted by several numeric variables; pH and other water chemistry measures, ; essentially abiotic variables that were measured for each site. 

While in the original data samples are not independent (some are taken from the same pond), I will use the water body worksheet that I made where a random site is selected, per temporal sampling session, for each water body. I don't really have the resources to do spatial awareness (or whatever it's called) analysis. Each water body has a code, could use that in model?

Or, for each water body per sampling session, I could take averages of the abiotic variables (eg take average of pH) and then total species richness for each water body and take presence/absence from there. Should talk to Mart about this except he's away rn. 

Or, for this project, I could stick to one water body--like Andover Pond, or Pymatuning, or state gameland 214, or something with interesting variation and at least three species and measure change across time in presence/absence.

I will wait to do presence of species predicting presence of other species for now because that will be too complicated without analyzing abiotic variables first.

Data Analysis Plan: Once data is made to be independent as best I can manage, it will be analyzed by logistic regression, eg multiple variables (PCA?) predicting the probability of a species being present or not.
> look at biostats_week12.Rmd for how to set up logistic regression

for once the data is all uploaded, 

sorting out data
```{r}
attach(X2019Duckweed_PresenceAbsenceData_v1)

#have to make many of these variables numeric
pH <- as.numeric(pH)
Conductivity <- as.numeric(Conductivity)
Dissolved_Oxygen <- as.numeric(Dissolved_Oxygen)
Densiometer <- as.numeric(`Densiometer_(CC %)`)
LM_cover <- as.character(LM_Cover)

```


pH analysis and plot for LM (Lemna minor)
```{r}
#read in data
#dw_presenceabsence <- read.table(X2019Duckweed_PresenceAbsenceData_v1)
#head(X2019Duckweed_PresenceAbsenceData_v1)


#scatterplot of data
plot(LM_Presence ~ pH)

#Fit a logistic regression
LM_presence_pH_Glm <- glm(LM_Presence ~ pH, family = binomial(link = logit))

#Add logistic regression curve to scatter plot
xpts <- seq(0, 14, 
	length.out = 100)
ypts <- predict(LM_presence_pH_Glm, newdata = data.frame(pH = xpts), 
	type = "response")
lines(xpts, ypts)

#Table of regression coefficients, with parameter estimates
summary(LM_presence_pH_Glm)

#95% confidence intervals for parameter estimates. It is necessary to load the MASS package first.
library(MASS)
confint(LM_presence_pH_Glm)

#Predict probability of presence (mean presence) for a given x-value, 10 pH, including standard error of prediction.
predict(LM_presence_pH_Glm, newdata = data.frame(pH = 10),
	type = "response", se.fit = TRUE)


#Analysis of deviance table, with a of the null hypothesis of zero slope
anova(LM_presence_pH_Glm, test = "Chi")
```

Spirodela presence/absence based on pH
```{r}
#scatterplot of data
plot(SP_Presence ~ pH)

#Fit a logistic regression
SP_presence_pH_Glm <- glm(SP_Presence ~ pH, family = binomial(link = logit))

#Add logistic regression curve to scatter plot
xpts <- seq(0, 14, 
	length.out = 100)
ypts <- predict(SP_presence_pH_Glm, newdata = data.frame(pH = xpts), 
	type = "response")
lines(xpts, ypts)


```



Make model predicting Species Presence/Absence based on many variables
```{r}
#Fit a regression
LM_presence_pH_Glm <- glm(LM_Presence ~ pH*Conductivity*Dissolved_Oxygen)
summary(LM_presence_pH_Glm)
#?glm

#Spirodela polyrhiza
SP_presence_pH_Glm <- glm(SP_Presence ~ pH*Conductivity*Dissolved_Oxygen)
summary(SP_presence_pH_Glm)


```


Andover pond model
conditional selection by Water_Body==OPRW_1
```{r}

```

